A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-08-25 code/supplemental_v2.Rmd
Minor update to supplemental
Remove st2 from supplemental pdf

**Supplementary Table 1. Cluster-specific enriched pathways (effect size >= 0.30).**
```{r enrich-pathway-in-clust, fig.width= 9}
## docx friendly
load("../data/moderate_enrich_prob.Rda")

st1 <- moderate_enrich_prob %>% filter(difference >= 0.30 & signif <= 2500) %>% ## filter effect size >= 0.20 and enrichment at a 95% CI
  dplyr::select(path, clust, signif, difference) %>% 'colnames<-' (c("Pathway", "Cluster", "Signif", "Enrichment_Score")) ## select relevant column and rename col

@@

![**Supplemental figure S6. Survival fit independent of disease status.** _The red and blue curves show the survival of people in the US by female and male, respectively. The orange and green curves show the model fits._]

>>>>

**Supplementary Table 2. Cluster-specific enriched pathways (effect size >= 0.30).**
```{r enrich-pathway-in-clust, fig.width= 9, eval = FALSE}
## docx friendly
load("../data/moderate_enrich_prob.Rda")

table_st2 <- moderate_enrich_prob %>% filter(difference >= 0.30 & signif <= 2500) %>% ## filter effect size >= 0.20 and enrichment at a 95% CI
  dplyr::select(path, clust, signif, difference) %>% 'colnames<-' (c("Pathway", "Cluster", "Signif", "Enrichment_Score")) ## select relevant column and rename col

@@

### 2020-08-25 text/manuscript.Rmd
Remove st2 from supplemental pdf

We investigated this dataset further using multiple correspondence analysis
(MCA), and visually summarized the analysis with UMAP (**figure 2**). We used
the resulting UMAP graph coordinates to perform density based clustering with
HDBSCAN, which resulted in identification of 10 well-defined clusters capturing
about 80% of the tumor samples. To capture the remaining samples into one of
these 10 clusters we used kNN (see **Supplementary Methods** for details on all
clustering algorithms).

@@

**Figure 3: Pathway-based clustering independent of tissue-of-origin** 
A) Sankey plot showing correspondence between cancer type and cluster identity.
B) Projection of cancer type identity onto UMAP. Abbrevs: BRCA = Breast cancer,
COAD = Colorectal adenocarcinoma, HNSC = Head and neck squamous cell, PAAD =
Pancreatic adenocarcinoma. C) Projection of breast cancer subtypes onto the
UMAP. D) Projection of BRCA1/2 somatic mutation onto the UMAP. E) Projection of
tumor stage onto the UMAP, regardless of cancer type. E) Projection of
metastatic status onto the UMAP. Abbrevs: M0 = non-metastatic tumors, M1 =
metastatic tumors.

@@

To discover what pathways are most important for clustering, we calculated
percent enrichment for each pathway _within cluster_ relative to all other
clusters combined (see methods) and ranked pathways for each cluster from
highest to lowest enrichment. We visualized the enrichment as a heatmap
(**figure 5B**). Using this approach, we identified 25 pathways highly enriched
(enrichment score $> 0.3$, see methods) in cluster 2, 9 pathways enriched in
cluster 3, 24 pathways in cluster 4, 10 pathways in cluster 5, 10 pathways in
cluster 6, 24 pathways in cluster 7, 24 pathways in cluster 8, 10 pathways in
cluster 9, and 0 pathways for clusters 1 and 10. Clusters 7, 8, and 9 had
several pathways in common. To explore the specific pathways marking each
cluster, we projected disruptions for each of the 377 pathways onto the UMAP
(**supplementary figure S4**). Clusters 3 and 5 were distinguished by metabolic
pathways including RNA and protein biosynthesis (**supplementary figure S4**).
Similarly, cluster 4 was distinguished by mutations affecting regulation of DNA
and histone methylation ("DNA methylation", "PRC2 methylates histones and DNA",
and "Nucleosome assembly"). Clusters 7-9 have in common mutations in
extracellular, intracellular, and immune-related signaling pathways (see
**figure 5C** and **supplementary figure S4**). Cluster 2 had the highest
pathway enrichment levels of the three, having mutations in hedgehog signaling,
"β-catenin degradation", "cellular response to hypoxia", "regulation of cell
cycle" and "apoptosis" among others.

>>>>

We investigated this dataset further using multiple correspondence analysis
(MCA) [@JSSv025i01], and visually summarized the analysis with UMAP (**figure 2**). We used
the resulting UMAP graph coordinates to perform density based clustering with
HDBSCAN, which resulted in identification of 10 well-defined clusters capturing
about 80% of the tumor samples. To capture the remaining samples into one of
these 10 clusters we used kNN (see **Supplementary Methods** for details on all
clustering algorithms).

@@

**Figure 3: Pathway-based clustering independent of tissue-of-origin** 
A) Sankey plot showing correspondence between cancer type and cluster identity.
B) Projection of breast cancer subtypes onto the UMAP. C) Projection of BRCA1/2
somatic mutation onto the UMAP. D) Projection of tumor stage onto the UMAP,
regardless of cancer type. E) Projection of metastatic status onto the UMAP.
Abbrevs: M0 = non-metastatic tumors, M1 = metastatic tumors.

@@

To discover what pathways are most important for clustering, we calculated
percent enrichment for each pathway _within cluster_ relative to all other
clusters combined (see methods) and ranked pathways for each cluster from
highest to lowest enrichment. We visualized the enrichment as a heatmap
(**figure 5B**). Using this approach, we identified 25 pathways highly enriched
(enrichment score > 0.3, see methods) in cluster 2, 9 pathways enriched in
cluster 3, 24 pathways in cluster 4, 10 pathways in cluster 5, 10 pathways in
cluster 6, 24 pathways in cluster 7, 24 pathways in cluster 8, 10 pathways in
cluster 9, and 0 pathways for clusters 1 and 10. Clusters 7, 8, and 9 had
several pathways in common. To explore the specific pathways marking each
cluster, we projected disruptions for each of the 377 pathways onto the UMAP
(**supplementary figure S4**). Clusters 3 and 5 were distinguished by metabolic
pathways including RNA and protein biosynthesis (**supplementary figure S4**).
Similarly, cluster 4 was distinguished by mutations affecting regulation of DNA
and histone methylation ("DNA methylation", "PRC2 methylates histones and DNA",
and "Nucleosome assembly"). Clusters 7-9 have in common mutations in
extracellular, intracellular, and immune-related signaling pathways (see
**figure 5C** and **supplementary figure S4**). Cluster 2 had the highest
pathway enrichment levels of the three, having mutations in hedgehog signaling,
"β-catenin degradation", "cellular response to hypoxia", "regulation of cell
cycle" and "apoptosis" among others.
